### 2023-10-25 Latex files/Elisevier/Cover_letter CSA.tex
Update to Latex files sub folder

\documentclass[]{letter}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000
\usepackage[absolute,overlay]{textpos}

\begin{document}

\begin{textblock*}{13cm}(4cm,2.54cm) % {block width} (coords) 
	\begin{letter}{Dear Editors,}
	\opening{}
	We are pleased to submit a manuscript titles \textbf{‘Investigation of Multi-Angle Detection Schemes for Light Scattering of Optically Trapped Asymmetric Colloidal Particles’} P Parthasarathi, M Haw, L Lue, J Sefcik,
	and myself for publication as an article in Optical Communications.

	This work introduces a new method to quantitatively characterize
	optically trapped colloidal particles, by measuring light scattered at
	multiple angles and analysing the data based on Bayesian inference.
	To demonstrate the method, we apply it estimate the instantaneous
	orientations of an optically trapped asymmetric colloidal dimer.  This
	technique can be easily extended to characterize the size and shape of
	optically trapped colloidal particles and offers a powerful tool to
	study colloidal systems. The ability to interpret light scattering signals from an optically trapped entity should be of interest to readers
	of your journal.

	We hope that you will consider our work for publication.  If you need
	any further information or materials, please let me know.  Thank you
	for your attention.

	Sincerely yours,

	Daniel Maciver \\

	Department of Chemical and Process Engineering \\
	University of Strathclyde \\
	James Weir Building, 75 Montrose Street \\
	Glasgow, G1 1XJ \\
	United Kingdom \\

	phone: +44-779-344-5441 \\
	fax:   +44-779-344-5441 \\
	email: daniel.maciver.2016@uni.strath.ac.uk
	\end{letter}
\end{textblock*}

%enclosure listing
%\encl{}

\end{document}