### 2024-07-24 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase separation 
used by humanity\cite{Schoen1956}, put simply, it is the 
formation and growth of a new structured phase within 
a disordered bulk phase. This has applications in a 
number of industries such as pharmaceuticals, food 
production, and electronics \cite{Myerson2002}; where 
the extraction of dilute materials can help improve 
product quality while keeping production costs low. 
The industrialisation of crystallisation has allowed 
engineers to reliably and efficiently induce the crystal 
formation within a bulk phase. However, while on a 
large scale crystallisation is seemingly an understood 
physical process, just a small amount of investigation 
into the literature reveals that at a micro scale there 
is yet to be unifying theory that can accurately explain 
the process of crystallisation \cite{Fu2021}. Of which 
there two primary areas of focus, nucleation and crystal 
growth. The latter focusing on how an already stable 
crystal grows and how it takes on its final shape; 
whereas the former is more concerned with which factors 
contribute to its formation in order to control crystal growth.

The focus of this research is partly based around investigating
the impact of a focused electromagnetic field can influence the 
nucleation process by inducing shear flow within the bulk phase. 
This in hope that the localised EM field can be used to study 
different phenomena in nucleation in order to develop a better
understanding of the underlying mechanisms.

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain in equilibrium while 
below a specific concentration ($C_{eq}$) - below which 
the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate 
the two phases. Once $C_{eq}$ is exceeded there is a chemical
 potential difference driving the solution to separate the 
two phases. Since different combinations of solute and solvent
will have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by using 
'supersaturation', one form of measuring supersaturation is shown 
below \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
is the equilibrium concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where short and consistent 
residence times are desirable. At a small scale one can estimate the 
nucleation rate by making repeated measurements of sample solutions 
and seeing how many have nucleated after a given time, giving us a 
Poisson distribution.

\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$
is the volume of the individual samples, and $M^*(t)$ \& $M$ are the 
number of nucleated samples and the total number of samples used respectively.
While this is useful for studying the effects of different parameters
at a small scale, for industrial applications there are so many
additional factors at play that can throw off this estimation.

In contrast, secondary nucleation is the result of a initial 
seed crystal inducing further nucleation within the bulk solution
\cite{Botsaris1976}, in absence of a seed crystal nucleation is
near impossible. The research of Secondary Nucleation is fascinating
for its own merits; is it speculated that the supersaturation 
barrier is non-existent \cite{Cashmore2022}, and furthermore there 
are a multitude of mechanisms that can induce secondary 
nucleation making it imperative that researchers can control the
nucleation rate at an industrial scale.

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by Agrawal and Paterson \cite{Agrawal2015}}
\end{figure}

\section{Nucleation Theories}
\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the 
original theory was first formed from the works of Volmer 
and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
While initially it was more focused into describing droplet 
formation in condensing vapours it was extrapolated to describe
crystallisation. The central premise of classical theory 
is that nucleation occurs stochastically throughout the bulk
phase due to collisions between individual solute molecules,
ions, or atoms; at the same time the bulk phase is resistant
to the formation of a new phase. The competition between
these random collisions and the bulk solution can be used to 
predict the probability of a newly formed nucleus.

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}

Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	     \label{fig:free_energy}
\end{figure}

The maximum value of $\Delta G_{tot}$ is the free energy barrier 
that any newly formed nucleus needs to overcome in order to stabilise.
The nucleation rate (the volume of new crystalline material formed per
unit time), is therefore commonly defined as being dependent on the 
energy barrier $\Delta G^*$:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}

Where $A$ is a pre-factor that can be fine tuned to the exact demands
of the system (typically involving Zeldovich Factor, $Z$) the free 
energy barrier can be found by finding the turning point of $\Delta G_{tot}$.

CNT is often regarded as a good description of the macro system, its obvious
that for all crystallization systems there is an inherent energy barrier that 
dictates the nucleation rate. Where it falters is in its predictive ability, 
both in estimating nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in 
the structure of newly formed nuclei \cite{Lee1999, Yau2001}. Recent studies 
suggest classical nucleation is merely one of many possible pathways that can 
be taken to produce a structured crystalline phase. Prompting the development of alternative theories to better describe the nucleation process.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, colloid simulations 
showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) 
interactions allow for the formation of a liquid-liquid metastable phase from 
which a new solid phase could form \cite{Anderson2002, Karthika2016}. 
\textit{in situ} techniques for studying nucleation several papers reported 
the presence of stable liquid-like clusters that formed prior to nucleation
\cite{Savage2009, Wolde1997, Soga1999}. It can be understood by Oswald's rule
\cite{Ostwald1897}, which says that any crystallising system does not immediately 
take the path to the lowest possible energy state but instead first transitions 
to the state with the smallest free energy barrier, further transitions can still 
occur but this pathway minimises the overall free energy cost. If we consider 
the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider 
a case where the supersaturation is low ($S\approx1.01$) then a dense liquid 
droplet will have a lower chemical potential than the surrounding fluid but 
the crystalline phase will have a subsequently lower chemical potential than 
the dense liquid phase. Initially it was suspected that the formation of the 
dense liquid phase was a result of stochastic density fluctuations in the system; 
further simulations suggest that the the local bond order is a stronger driving 
force than the local density \cite{Tan2013}.

The current research into two step nucleation (or now more commonly referred 
to as multi-step nucleation theory) is developing a robust framework to 
describe what nucleation pathway will occur given the initial conditions. 
Reviews of all currently documented nucleation pathways identified highlighted 
the fact the need for a the development of \textit{in situ} techniques that 
can induce nucleation locally but can also reliably identify the nucleation 
pathway across a broad range of experimental conditions \cite{Karthika2016, Fu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Crystallisation methods}
\subsection{Cooling Crystallisation}
For some binary mixtures the supersaturation is heavily dependent on 
the solution temperature, therefore a simple method of producing crystals 
is by cooling the mixture to induce crystal formation. At ambient 
temperatures the solute concentration is too high to be fully incorporated
into the solution ($S\gg 1$), after heating however the solute is fully dissolved
into the solution ($S<1$). Now as the mixture is allowed to cool to room
temperature the supersaturation will increase until crystal formation begins,
the rate of cooling drastically influencing the size and number of crystals
produced.

If $dT/dt$ is high then the final product will consist of large 
crystal and be low in number, as the nucleation rate is directly related
to the supersaturation only a handful of nuclei can form before the remaining
solute grows onto the surface. If $dT/dt$ is low then the final product will 
consist of smaller crystals and be far more numerous, as the supersaturation 
is so large that the new nuclei are forming continuously. Between these two 
extremes, one can define the meta-stable zone width, a region in the 
temperature-concentration phase space where both nucleation and crystal growth
can be reliably controlled. The lower limit being given by the solubility curve
of the binary mixture, and the upper limit being defined by the metastable zone.

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{MSZW.png}
	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red). 3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
	\label{fig:MSZW}
\end{figure}

The viability of cooling crystallisation is dependent on the meta-stable 
zone width, too narrow and the process is difficult to control, too wide 
and the crystal growth rate may be insufficient for the desired outcome.

\subsection{Evaporative Crystallisation}
In situations where control of the final product size or shape is not the main
focus, often the cheapest method of producing a crystalline product is simply
to allow the solvent evaporate and separate from the solute. Depending on the 
total volume of solvent to evaporate this process can take on time scale of 
several days to complete.

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since the 1970s
when Ashkin \cite{Ashkin1970} first showed that focused light was capable 
of trapping micron sized particles due to light exerting 'radiation pressure'. 
The working principle was that a light source such as a laser could trap 
small objects within a 2D plane, as long as the light source had an 
approximately Gaussian profile. Soon after, Ashkin showed that the introduction 
of a microscope objective would allow one to focus the light source to a 
diffraction limited point that would stably trap small objects within a 
confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study 
biological material and would later be used to probe microscopic properties 
such as the formation of colloidal aggregates \cite{Yi2021} to the drag 
forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the 
predictable behaviour of light, optical tweezers have become essential for 
measuring and exerting precise forces on the magnitude of pico-newtons 
allowing one to probe the material properties of the smallest materials.

\subsection{Literature related to laser induced nucleation}
From as early as 1996 it has been known that laser irradiation is a viable
method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, 
the first reported case was notable as it used a 1.064 $\mu m$ laser meaning 
it was unlikely to be a photo-chemical reaction but rather a physical one. 
Future research has found nucleation can be induced by 1 of 3 routes involving 
direct laser induction: firstly, Non-photochemical laser induced nucleation 
(NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996,
Garetz2002,Sun2006}, several papers have debated the exact mechanism that 
induces NPLIN \cite{Garetz2002, Knott2011}. Two suggested hypothesis are: 
an optical Kerr effect, where the solute molecules are aligned to lower the 
nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in 
which solute clusters are stabilised within the electric field which drastically 
increases the likelihood of nucleation\cite{Alexander2008}. Both hypothesise 
have their limitations and there has yet to be a single theory that explains 
NLPIN thoroughly. In either case the mean pulse intensity needs to be kept 
relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses 
lead to a completely different nucleation mechanism.

High intensity laser induced nucleation (HILIN), where the pulse intensity 
is on the order of several $PW/cm^2$ is far simpler a mechanism to explain 
in comparison to NPLIN. The production of nuclei can be wholly associated 
to a cavitation process within the target solution, where the laser focus 
results in thermo-cavitation and the subsequent pressure change leads to
a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}. 
There is still not a general consensus on how cavitation influences the 
local supersaturation \textcolor{red}{[ , ]}, nor is there a clear 
understanding of what properties of the crystal can be controlled 
\textcolor{red}{[ , ]}. It has been suggested that in theory any solution 
can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a 
strong understanding of the phenomena both before and after cavitation 
occurs.

Lastly, there is trapping induced nucleation, this is where optical tweezers 
come into play, the optical trap has been shown to have different effects 
on supersaturated solutions depending on where it has been focused. When 
focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ 
were shown to create a dense liquid droplet of glycine and water 
\textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid 
region was populated by clusters that would consolidate together upon being 
focused by the optical trap \textcolor{red}{[ ]}. Molecular simulations of 
glycine solutions showed that these clusters are unstable when using pure 
glycine below the saturation point \textcolor{red}{[ ]} suggesting that the 
clusters are formed due to glycine reaction products. When the optical trap 
was moved from the cover slip to the air-solution interface nucleation would 
occur before a dense liquid region could form \textcolor{red}{[ ]}. Repeated 
experiments where the laser is focused on the air-solution interface have 
lead to a variety of different nucleation events. In some instances the 
nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; 
whereas allowing a solution to age results in the formation of amorphous 
precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}. 
The precursors are only seen when the solution is irradiated by an optical tweezer 
and the growth rate can be controlled somewhat by varying the laser power 
\textcolor{red}{[ ]}. Notably the only work has been done with simply 
irradiating the solution with a trapping potential, there has not been an 
attempt to introduce a trappable object into the solution, the trapping 
potential has been used to influence the growth of a crystal front 
\textcolor{red}{[ ]}.

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and 
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

Spin angular momentum (SAM) is attributed to the spin density of the field, 
early research has shown that the spin density is non-zero for any beam despite
the fact that the total SAM transferred to a medium is 0 \cite{Svak2018, 
Bliokh2014}. This has sparked debate if SAM is even a physical quantity as 
it does not aid in the transport of energy directly \cite{Bliokh2014} and 
so cannot be directly observed in some cases despite being non-zero. This 
paradox is resolved by representing the wave as an array of spin momentum 
loops that all together cancel one-another out when the medium is homogeneous.
Spacial inhomogeneities cause these spin loops to non longer be equal, resulting
in non-zero spin density, anisotropic mediums (such as birefringent crystal 
lattices) experience a transfer of spin angular momentum allowing for optical
torque transfer. It has even been shown that in the case of homogeneous spheres, 
the spin component of a focused beam can result in additional orbital angular
momentum via spin-to-orbital conversion \cite{Yevick2017}.

Birefringence is a material property often seen in crystalline materials, 
if the crystal lattice has a different structures when viewed at different 
orientations then light will be refracted differently depending on its 
polarisation. For circularly polarised light the inhomogeneity results in 
a high degree of SAM being transferred to the target object \cite{Parkin2009,
Arita2016}, this has been exploited to rotate microspheres as fast as 1000 Hz 
while suspended in a bulk medium \cite{Arita2016} as well as a means of 
measuring the local temperature and shear response of said medium 
\cite{Millen2014, RodriguezSevilla2018}. Calculating the optical torque 
applied to a birefringent material is given via:

\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\theta$ is the angle between the particle's long axis and the Poynting
vector of the local EM field, and $\phi$ is the phase shift in the EM field.  
The first term represents the 'orientational' torque which is due to the target 
particle being aligned along the EM field, when aligned $\theta=0$ meaning 
the entire term is negligible for particle's with a stable orientation. The 
second term is due purely to the polarisation of the optical trap, for 
circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred 
to the target particle. Birefringence can also be induced if the target 
particle has an anisotropic shape, in particular if the particle shape is 
elongated along one major axis; the go to particle shape is a silica dimer 
(two spheres tangentially attached) due to silica's stability and strong 
adhesion. Using a silica dimer research groups have achieved a rotation 
frequency in the realm of several GHz \cite{Ahn2018} in a vacuum.

\subsection{Shear induced Nucleation}
It has long been known that fluid shear rate plays a role in influencing
nucleation; however, the exact relationship between shear rate and nucleation
rate has only been recently understood for specific solutions. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

\section{Aims and Objectives}
Overall the aim of the PhD is to study viability of using micro-rotors to
generate localised fluid flow around the beam focus. The results are reported
in chapter 3, this is then succeeded by experimental work where we 
instead using a galvano-mirror to generate shear flow. While overall unsuccessful
the addition of a moving beam focus showed that the growth of a nucleus can be localised around the trap focus.

In addition, computer simulations studying the dynamics of asymmetric dimers 
and their applications as complex micro-rotors are discussed in chapter 4. 
Lastly two novel methods of measuring rotational motion are discussed in 
chapter 5; firstly via a novel detection fibre method that allows for 
instantaneous measurements of the orientational behaviour of optically 
trapped ellipsoids/dimers; and secondly we create a simulative quadrant 
photo diode that replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to measure the 
optical torque applied to a non-birefringent particle.