### 2024-09-16 Latex_files/Thesis/chapters/chapter 2/main.tex

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the differences in 
scattering theories, and the broader theoretical calculations used
throughout the rest of the Thesis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

Proper understanding of optical tweezing requires an understanding of
how the trapped particle interacts with the trapping laser. From an 
electromagnetism perspective the laser creates a spatially and 
temporally coherent electric field that scatters light off of a 
trapped particle.  The laws governing electric and magnetic fields 
are summarised most succinctly via the Maxwell equations. The 
differential forms of which are given below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, and $\epsilon_0$ is the permittivity of free space.

These four equations describe how the electric and magnetic fields behave
and relate to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
The latter two are defined by:
\begin{align}
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ is the 
magnetisation field, these fields arise due to the bound charges 
throughout the medium interacting with the EM field.

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a 
conservative force brought about by the polarisation of dielectric 
materials which is directed towards the point of maximum intensity 
(for a simple Gaussian beam this would be at it's focal point) \cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
higher numerical aperture lenses in order to increase the intensity, 
thus allowing for stronger gradient forces at the trade off of 
decreased trapping depth.

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far exceeds the both the scattering force, 
and the thermal energy of the particle.

\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Shown below is a typical 
force displacement curve for a single sphere in a focused Gaussian beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both 
the laser power and the contrast in refractive indices between the 
medium and the particle. Fully describing and calculating the 
optical forces is dependent on the ratio between the particle size 
and the trapping wavelength. Depending on this ratio, the optical 
forces can be described differently.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, this theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam, one coming close to the centre, and the other ray 
coming from the edge. As each ray is refracted by the target sphere a 
force is imparted onto it, the total force imparted is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lornez-Mie Theory}

The Lorenz-Mie theory provides an exact solution to the Maxwell 
equations for the scattering caused by an isotropic sphere. This 
theory describes the scattered wave given off by a dielectric 
sphere when incident by a plane wave as a summation of partial 
spherical waves. For any spherical wave the vector fields must 
solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves, both of which can be converted between 
with relative ease. For example, a plane wave electric field can 
be expanded into spherical harmonics and likewise any spherical 
wave can be described as a combination of plane waves offset from 
one another. For a single plane wave the incident, internal, and 
scattered fields are given as: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  \label{eq:scattered}
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M_{nm}$ 
and $N_{nm}$ are the vector spherical harmonics. The superscripts
denote whether the field is inward or outward propagating; so a 
superscript (1) is inward propagating, and (2) is outward propagating.  
For the incident field computing its expansion coefficients is possible 
via analytical methods and are completely dependent on the beam 
conditions imposed by the user. However, computing the expansion 
coefficients of the internal and scattered fields is far more 
tedious depending on the shape of the target and what properties 
of the scattered field are desired --- see sec.~\ref{sec:scattering}.  
From an optical trapping perspective the force exerted by a focused 
electric field can be found by computing the Maxwell stress tensor 
which only requires the total magnitude of the incident and scattered 
fields. An alternative method is to compute the momentum carried by 
the incident and scattered beams and taking the difference between 
to measure the momentum transferred to the particle. This project 
utilises the results from Farsund \cite{Farsund1996} but in the 
form used by Crichton \cite{Crichton2000THEMD}. As an example the 
linear momentum carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

And the linear momentum carried by the scattered beam is given by:
\begin{equation}
	\label{eq:linear_momentum_out}
	\begin{split}
		\langle P_{z,\ out} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
\end{align}

Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in \ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the light being scattered - the theory describing the scattered 
light is better approximated by alternative theories. For 
particles that are far larger than the wavelength of light the 
scattering is best described by the Ray-Optics theory. In the 
opposite case the target particle is better approximated as a 
single electric dipole as described by the Rayleigh scattering 
field.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rayleigh Scattering}

The Rayleigh approximation is the limiting approximation for 
describing a particles scattering in an electromagnetic field 
whose wavelength is several times greater than the particle's 
size. The underlying theory is that a dielectric sphere can be 
treated as a dipole while in the presence of the electromagnetic 
field. In which case the scattering force is given simply by the 
scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al 
\cite{YasuhiroHarada1996} in MKS units as a restorative rectangular 
force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

Where $m$ is the relative refractive index ($n_1/n_2$), $\omega_0$ is 
the beam waist, and $a$ is the radius of the particle. The scattering force 
however, is dependent on the effective scattering cross-sectional area. 
\begin{align}
  F_{\rm scat}
  &= \hat{z} \left(\frac{n_2}{2}\right) C_{pr} I(r) \\
  \text{where:} \nonumber \\
  C_{pr} &= \frac{8}{3} \pi (ka)^4 a^2 \left(\frac{m^2-1}{m^2+2}\right)^2
\end{align}

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole. Unlike Lorenz-Mie scattering, there is no need to compute the 
scattered field (see \eqref{eq:scattered}) for an individual sphere. 
Higher complexity scattering problems can be simplified by subdividing 
the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
in which case the scattered field is now necessary. As the target 
particle gets larger this assumption fails to accurately describe 
the trapping force. For particle sizes close to the laser wavelength 
the trapping force is best described by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle. While one can compute this by directly applying
Lorenz-Mie theory, describing the interactions between focused beams and 
complicated shapes can be computationally slow and memory intensive. As a
result, several alternative scattering approaches have been developed that
allow for faster computation times simplifying the scattering problem.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes  
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface. These currents are chosen so that the
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. This choice of currents negates the need to 
compute the internal fields which reduces the scattering problem 
to a system of linear equations relating the incident beam coefficients 
to the scattered beam coefficients. The relationship between each can 
be summarised as:
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, 
and $a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. 
For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}). Developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered 
field from each sphere within the target cluster and the incident 
field itself. The scattering between spheres slowly converges to a 
final result but modern codes truncate the calculations to an upper 
number of iterations (denoted by $N_{max}$) to fit a desired accuracy 
and computational time.

The $T$-matrix method is exceptionally useful for computing the 
scattering from any arbitrary spherical aggregate. However, the 
$T$-matrix method by itself can be computationally taxing as the 
number of spheres increases \cite{Mackowski2011}. While it is 
possible to solve for the full electromagnetic field of each the 
entire cluster this is only applicable for a single orientation 
and can be even slower for large aggregates \cite{Mackowski1996, 
Xu1995}. The benefit of \textit{mstm} is that the major scattering 
properties (scattering and extinction cross sections, scattering 
matrices, etc) can all be computed both for single orientations, 
or averaged over multiple orientations to determine the average 
scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that can 
be applied to the scattering from particles of arbitrary composition
and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
the DDA method approximates the particle as being constructed of dipoles.
Each dipole interacts with both the incident field and the scattered fields
from every other dipole surrounding it. The resulting scattered field is 
identical to the scattered field produced by direct integration of 
Eq.~\eqref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
The integral form for the internal electric field inside a scatterer is 
given as:
\begin{align}
	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
defines the impulse response between two separate dipoles; and $\chi$ is the 
susceptibility of the medium, which describes the degree of polarisation of
the medium in the presence of an electric field. With the internal field 
calculated, the scattered field can be computed, one of the primary advantages
of DDA over the T-matrix method is that the composition of the target can 
be changed freely. When comparing different computational scattering methods,
the ECBM was found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the target's symmetry to speed up calculations \cite{Wriedt1998}. But when dealing with inhomogeneous media DDA is 
more efficient compared to ECBM.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules
diffusive behaviour, for the case of optical tweezers the most complete model 
of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
derivations are sufficient for macroscopic behaviours the Langevin equation 
better describes the microscopic characteristics of any diffusive behaviour.

The Fickian model describes the net flux of solute molecules into a finite
volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$
\cite{Gillespie2012}. It assumes that the solute molecules do not collide with 
one another or with molecules in the solution; overall, the Fickian model is
used to describe how solute molecules disperse over long periods of time. It 
does not provide any incite into the forces acting on individual molecules, 
whereas the Langevin model captures the physical interactions between an 
individual molecule and the surrounding fluid over a wide range of time scales.

The Einstein model of diffusion expands upon the Fickian model by considering
the collisions between individual molecules. If we consider a single particle 
suspended in a solution it will experience multiple collisions with other 
molecules \cite{Gillespie2012a} over a given time interval $\delta t$. The 
Einstein molecule allows us to consider how individual molecules behave while
suspended in a surrounding fluid. Where it begins to falter is when we consider
bringing $\delta t$ to $0$; in this scenario the molecule experiences not many
but only individual collisions. The Einstein model does not consider the inertia 
of the molecule in question and so for very short time frames the model implies 
that the molecule's velocity changes instantly after each collision \cite{Gillespie2012a,
Gillespie2012b}. Furthermore, the kinetic energy of each collision is not limited
by the thermal energy of the system, meaning that using the Einstein model to 
predict a particle's trajectory to an finite degree of accuracy implies that 
the particle has infinite kinetic energy \cite{Gillespie2012b}. This failure to 
describe motion over smaller time frames is addressed in the Langevin model by 
accounting for the fluid drag of the system, where any sudden change in velocity 
must result in a proportionally opposed drag force \cite{Gillespie2012c}.

The Langevin model of diffusion assumes that the net force on a particular 
particle is described fully by these individual collisions \cite{Gillespie2012c}. 
Unlike the Fickian model it provides a full description of the interactions 
between the target molecule and the surrounding fluid; while also being able
to describe its motion over any time scale, unlike the Einstein model:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}

Where the first term accounts for inertial forces, the second term 
accounts for friction forces which counteract the particles current 
motion ($\gamma_0$ is the friction coefficient), and the final term 
accounts for the random Brownian force. The $F(t)$ is there for 
convention which accounts for any external forces acting on the 
particle. These three terms are equal to the random Brownian motion 
of the surrounding fluid. We can say that the noise term $W(t)$ has 
a normal distribution, being scaled by the thermal energy of the 
system, with a correlation function of:
\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

The Langevin model can be extrapolated to describe the diffusive behaviour 
of an overall system, but for this project we can instead consider the 
behaviour of some particle with a diffusion tensor $D$ suspended in a 
viscous fluid and spacially localised by an optical potential with trap 
strength $\kappa$. Assuming the only external force acting on our particle 
is the laser, the net force should be exactly equal to force of the stochastic
collisions due to the fluids thermal energy. If we focus our analysis when 
the particle is stably trapped and assume that the trap is harmonic, we can 
model the trapping force as a Hookean spring ($F(t) \approx \kappa x(t)$). 
The full Langevin equation for an optically trapped particle is therefore given as:
\begin{align}
	\label{eq:langevin}
	m\frac{\delta^2x(t)}{\delta t^2} + \gamma_0 \frac{\delta x(t)}{\delta t} + \kappa_x x(t) = \sqrt{2k_BT}\eta(t)
\end{align}

\eqref{eq:langevin} provides an accurate description of strongly trapped 
particles. Despite this the analytical solution of the Langevin equation requires 
integration of the noise term making it difficult to simulate the 
trajectory of a given particle \cite{Volpe2013}. Instead, it is often far 
easier to solve the equation numerically and use the analytical 
solution to calibrate and extract information about the particle and fluid, 
and how the two interact with the optical trap.

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial 
elements in order to approximate the higher order terms. If we assume that 
$x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then 
we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}

Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
term n and the actual expression. If we limit our approach to the first 
derivative only, we find that for sufficiently small values of $R_1$ the 
velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}

By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps, we can 
discretise the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
noise, as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our noise as a random walk, where at each new time 
step, the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}

Where $w_i$ is a random real number between -1 and 1, we can say that 
it is normally distributed around 0 for simulation purposes. We can 
rearrange this for $x_i$ to approximate the Brownian motion of a 
particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
Now in the case of an optical trap, the restoration time scale is 
given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
is far greater than the characteristic time. Therefore for simulation 
purposes, we can neglect the particle's inertia which allows the motion 
of an optically trapped particle to be written as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} - \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

This result can be generalised for a 3-dimensional description of an 
optically trapped particle, where each Cartesian direction has its 
own unique characteristic restoration time. We see from the result 
that trajectory is dependent on only a handful of factors, the trap 
stiffness $\kappa_x$, the fluid viscosity $\gamma$, and the thermal 
energy of the system $k_BT$ (with the latter two being related by 
Einstein's formulation of the diffusion coefficient $D = k_BT/\gamma$). 
Therefore, by calculating these parameters to a high degree of precision 
allows one to get a precise description of the forces experienced by a 
target particle, which in the past has been used for highly accurate 
force transduction \cite{BergSoerensen2004, Smith2003}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.67\linewidth]{finite_differences.png}
	\caption{Example trajectory created using Finite Differences method
			 for a $2\mu m$ diameter sphere. Trap stiffness's were estimated using \textit{ott} at $\kappa_x = \kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$. The particle's motion can be localised around the shaded ellipsoid.}
\end{figure}

\section{Calibration Techniques}
\label{sec:calibration_techniques}
There are several approaches for calibrating and characterizing the 
optical trap, each approach has its drawbacks and benefits so each 
option should be chosen based on what elements want to be characterized. 
The basis for each of these methods stems from the analytical solution 
of the Langevin equation:
\begin{align}
	\label{eq:anylitical_lang}
	x(t) = x(0)e^{-t/\tau_{OT}}+\sqrt{2D}\int^t_0dsW_x(s)e^{-(t-s)/\tau_{OT}}
\end{align}

Positional data is often acquired in an experimental setting either using
image analysis or photodiodes to infer the particles position relative to
the beam focus. The former is often used in cases when precision is not a 
key concern, as often a standard CCD camera will have limited spacial and 
temporal resolution. The latter method often requires the use of a quadrant
photo diode(QPD), the data provided by the QPD will often need to be converted
to physical units to compute the force directly - see Eq.~\eqref{eq:conversion_factor}.

\subsection{Potential Well Analysis}
The Langevin equation for an optically trapped assumes that the trap 
acts similar to a Hookean spring that creates a potential well about 
its centre. Therefore a simple analysis method is to understand the 
height and width of said potential well.

Potential analysis is a useful technique for estimating the strength 
of an optical trap; this method assumes that the force acting on the 
particle is purely conservative, an accurate presupposition if we 
ignore the motion of the particle as it enters the trap. This is 
because the scattering force is far more significant far away from 
the potential well and is negligible if the trap strength is much 
greater than the thermal fluctuations. With this in mind we can 
write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{B}T}} 
\end{align}

which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}

Now assuming our laser acts as a Gaussian beam we should be able 
to describe the probability distribution $\rho(x)$ centred at 
some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}

By inserting \eqref{eq:prob_dist} into \eqref{eq:potential_well} 
we can fit the potential well in order to determine the trap strength 
$\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. 
This has some limitations in that the large fluctuations can throw 
off the fit meaning a longer acquisition time is necessary to properly 
fit the potential well, making it difficult to characterise weakly 
trapped particles who may not remain trapped for long. It also 
provides no information on the particle itself (i.e. the friction 
coefficient $\gamma$ and diffusion coefficient $D$).

\subsection{Equipartition method}
The Equipartition method is by far the fastest and simplest means 
for estimating the trap strength but unlike Potential Analysis is 
limited strictly to harmonic potentials. This can be often not the 
case for highly focused beams, as the trap strength can vary due to 
polarisation differences. Simply put we can use the equipartition 
theorem to relate the potential well to the particle's thermal 
energy using \eqref{eq:prob_dist}:
\begin{align}
	\left<U(x)\right> = \frac{1}{2}\kappa_x\left<(x-x_{eq})^2\right> &= \int_{-\infty}^{\infty}\rho(x)(x-x_{eq})^2 = \frac{1}{2}k_BT \\
	\implies \kappa_x &= \frac{k_BT}{\left<(x-x_{eq})^2\right>} 
\end{align}

By taking a time average over multiple trajectories to get 
$\left<x-x_{eq}\right>$ we can get a fairly accurate estimation 
of the trap strength. Because of this requires a time average of 
the particle's displacement any large errors in the position 
measurement can have knock-on effects. Likewise with the potential 
analysis route, no information is gleaned about the particle itself.

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing 
the random motion of a given particle (or group of particles). 
This is useful information if say for example we want to understand 
reaction kinetics on the surface of a catalyst, if we know how 
far its likely to move from the surface we can tell if its likely 
to react when a catalytic site becomes available. As it pertains 
to colloids, consider a suspension of silica spheres immersed in 
a fluid undergoing Brownian motion (as described by the Langevin 
Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}

Where $\gamma$ is the objects friction coefficient which for spheres 
is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white 
noise variable that is directly related to the thermal energy of 
the surrounding fluid. If the motion is truly random then we should 
see an average displacement of 0 regardless of how long we measure 
for. If we wish to understand the effects of a given external factor 
(i.e. an electric field or localised heating), simply looking at 
displacement will reveal nothing of value as its difficult to 
differentiate between diffusive and a biased motion.

For each sphere we can record its position in the $x-y$ plane and 
measure its displacement from a set reference point; for example 
with an optical tweezer this could be the beam focus. We can measure 
the MSD by forming a 'window' between two points in time of the 
trajectory (i.e. $t \&  t+\tau$) and sliding this window along the 
entire trajectory length - to eliminate -ve displacements we take 
the square - we can then take the average of this series. Repeating 
over a range of time lags allows us to describe the MSD as a function 
of $\tau$:
\begin{align}
	MSD(\tau) = \left<|x(t+\tau) - x(t)|^2\right>
\end{align}

If we use \eqref{eq:anylitical_lang} for an optical tweezer we can 
expand out the squared term to get an analytical expression for the 
MSD as a function of time lags:
\begin{align}
	\label{eq:MSD}
	MSD(\tau) = \left<|x(t+\tau)^2-2x(t+\tau)x(t)+x(t)^2|\right> = \frac{2k_BT}{\kappa_x}\left[1-e^{-\frac{\tau}{\tau_{OT}}}\right]
\end{align}

From this expression its evident that the mean squared displacement 
increases with larger values of $\tau$ until it reaches a maximum 
value as shown below by the dotted line.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{MSD.png}
	\caption{Example mean squared displacement using \eqref{eq:MSD}, for a $1\mu m$ sphere trapped by an optical potential well. The dotted line represents the upper limit of the sphere's displacement due to the optical trap.}
\end{figure}

The MSD plot can be subdivided into two regimes, when $\tau \gg \tau_{OT}$ 
the particle is experiencing the harmonic potential described by 
the equipartition theorem, and when $\tau \ll \tau{OT}$ the particle 
is said to be freely diffusing within the trap focus. Of course for 
a freely diffusing object the MSD will never reach a plateau value, 
comparing MSD's for different particles provides a simple visual 
indicator of the difference in trapping strength. The MSD method is 
an already very versatile analytical tool for diffusive motion, however 
it is rather slow in computing time meaning it is only really beneficial 
when a high degree of accuracy is required and shorter time resolutions 
are unavailable - such as using a quadrant photo diode instead or a high 
speed CCD.

\subsubsection{Angular Mean Square Displacement (MSAD)}
It is also possible to plot the angular MSD (MSAD) using simulative data, 
however using this to describe the torque acting on an asymmetric object 
is difficult due to fact there is no simple approximation of the rotational
stiffness. Vigilante \textit{et al} \cite{Vigilante2020} derived the upper 
limit of a dimer's MSAD along its long axis by assuming it was strongly 
trapped and so had limited angular motion, there expression gives:
\begin{align}
	\lim_{\tau\to\infty}\left<(\Delta u_z)^2\right> = 
	2\left[1-\frac{1}{4\beta\kappa_r} 
	\left(\frac{exp(\beta\kappa_r)-1}
	{exp(\beta\kappa_r)F(\sqrt{\beta\kappa_r})
	}\right)^2\right]
\end{align}

Where, $u_z$ is the unit vector connecting the two spheres of the dimer, 
$\beta=1/k_BT$, $\kappa_r$ is the rotational stiffness of the trap, and $F$ 
is Dawson's integral \cite{Oldham2008}. Vigilante's paper expressed that 
they couldn't compute $MSD(\tau)$ because they couldn't solve the 
Einstein-Smoluchowski equation which describes the diffusion constant for 
dielectric particles. This would require a full description of a 
particle's electrical mobility and charge distribution - the latter could 
be achieved via a discrete dipole approximation, the former would be 
dependent on both the particle's position and relative orientation to the 
electric field.

\subsection{Power Spectrum Density (PSD)}
The power spectral density (PSD) method is by far the most versatile 
method for observing the dynamics of any object within an optical trap, 
allowing for fast calibration times while also quickly filtering out 
typical noise sources. Taking the Fourier transform of a particle's 
trajectory yields:
\begin{align}
	\hat{x}(f) = \frac{(2D)^{1/2}\hat{\eta}}{2\pi(f_c-if)}
\end{align}

where $\hat{\eta}$ is the Fourier transform of the white noise, 
where the values are exponentially distributed as opposed to being 
normally distributed in the time domain \cite{BergSoerensen2004}, 
the correlation function is given as:
\begin{align}
	\left<\hat{\eta_k}\hat{\eta_l^*}\right> = t_{msr} \delta_{k,l} \rightarrow \left< \eta^4 \right> = 2t_{msr}
\end{align}

We can therefore ignore the white noise from our analysis by looking 
at the spectral density of $\hat{x}(f)$: 
\begin{align}
	\label{eq:lorentzian}
	S_x = \frac{\hat{x}^2}{t_{msr}} = \frac{D}{2\pi(f_c^2+f^2)}
\end{align}

Where $t_msr$ is the period over which we collected data. The Lorentzian 
can be fitted via a simplified geometric series $S_x = 1/(A+Bf_k^2)$ 
which allows us to compute both the diffusion coefficient (in arbitrary 
units) and the corner frequency $f_c$ which is directly related to 
the trap strength via $f_c = \kappa_x/(2\pi\gamma)$. The Lorentzian shape
implies that the trap is harmonic - but not symmetric - which assumes that 
the particle itself is an isotropic scatterer. Anisotropic scatterers will
not produce a Lorentzian curve due to the coupled rotation and translation, 
though depending on the nature of that rotation (i.e. whether it is stochastic
\cite{Bang2020} or periodic \cite{Yogesha2012}) will effect how drastically it 
departs from a typical Lorentzian curve.

Like with the analytical expression of the mean squared displacement we 
see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value 
that when converted to length units represents the maximum displacement 
the particle can move beyond the focus, however when $f\gg f_c$ the 
PSD falls off exponentially which denotes the particle is freely diffusing 
within the beam focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{PSD.png}
	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water. The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector. Using a correction factor (see \eqref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
\end{figure}

The Lorentzian relationship is only valid for frequency terms up to 
the Nyquist frequency (half of our sampling rate), this is because we 
are only taking a finite sampling of the particle's trajectory meaning 
the signal is aliased. Berg and Sorensen provide a suitable modified 
Lorentzian to account for the aliasing effects \cite{BergSoerensen2004}:
\begin{align}
	\label{eq:alaised_lorentzian}
	S_x = \frac{(\Delta x)^2\Delta t}{1+c^2-2c\cos{2\pi f_k\Delta t/N}}
\end{align}

Where N is the total number of samples taken, $\Delta x \ \& \ c$ 
have no direct physical interpretation and are defined in 
\cite{BergSoerensen2004}. Further modifications can be made to the 
power spectrum model but this is only useful when a high degree of 
accuracy is necessary. Typically power spectra are recorded using 
a Quadrant Photo Diode (QPD), which records motion in voltage units, 
not in units of length. There are two main methods for converting to
physical units: If multiple photodetectors are available then a 
differential interference contrast (DIC) system can be used to compute
the linear relationship between the beads displacement and the photodiodes
signal \cite{Capitanio2002}. While this is useful when high force precision 
is necessary it also allows you to collect information about the particle's 
motion along a single direction, making it ideal for less focused beams. 
Alternatively, if the size distribution is very wide but each particle can be 
accurately sized, then a conversion factor can be approximated by comparing 
the fitted value of the diffusion coefficient, and the reported value 
given by the Stokes-Einstein relation.
\begin{align}
	\label{eq:conversion_factor}
	D_{SE} = \frac{k_BT}{\gamma_0} \Rightarrow Conversion\ Factor \ [m/V]= \sqrt{\frac{D_{SE}}{D_{fit}}}
\end{align}

With the latter method, the local fluid viscosity must be known to a 
high degree of accuracy, depending on the local heating effect this 
may be as trivial increase or it may be significant enough to drastically 
alter the characterisation. PSD analysis is often seen as the gold 
standard for calibration as it can be fine tuned to the point that 
optical forces can be computed on the order of $10^{-15} N$ 
\cite{BergSoerensen2004}, it captures all of the information acquired 
by other calibration techniques while filtering out noise and requiring 
a relatively small amount of data collected.

\subsubsection{Power spectral analysis for rotating objects}
The scattering from rotating objects can be partially characterised via
the power spectrum density method. Typically papers reporting on the 
rotation rate of a trapped object will collect a power spectrum and look
at the maximum frequency term to determine its rotation rate. However this
often neglects any information on the trapping forces acting on particle
and only looks at the torque applying around one of the particle's primary 
axis. There has only been one notable effort to fully characterise the 
full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
a theoretical model for that relates the power spectrum to both the 
rotational and translational motion simultaneously.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}
\label{sec:simulations}
Later chapters cover the dynamics of spherical aggregates and anisotropic 
scatterers, these subjects are particularly difficult to characterise 
using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
As an example consider a symmetric dimer as a paradigmatic aggregate; 
if we consider the Langevin equation for such a aggregate within an optical 
trap we have:
\begin{align}
\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}

Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
motion is generalised to a 3 dimensional case. Except now, the dimer 
is undergoing random rotational motion in addition to its Brownian 
translational motion the first term on the right hand side is no longer 
purely a function of the dimer's position but also on its orientation. 
The rotational form of the Langevin equation for a dipole within an external 
potential is given as:
\begin{align}
  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

Where $\vec{u}(t)$ is the unit vector aligned along the centres of the 
two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational 
friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a 
sphere, if the dimer is within a harmonic potential we can write the first 
term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times 
\vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. 
$\lambda(t)$ is the Brownian rotations from the surrounding fluid, 
like in the translational case the Brownian rotations are normally 
distributed and are also uncorrelated so that:
\begin{align}
  \left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
\end{align}

For an asymmetric scatterer whose radius is comparable to that of the 
electric field's wavelength we now have a system of simultaneous equations:
\begin{align}
	\label{eq:full_langevin}
  \frac{{d}\vec{r}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
  \\
  \frac{{d}\vec{u}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

Fortunately, we do not need to solve these directly as the latter two
random variables can be easily approximated if the thermal energy of
the system is known, and the rate of change can be assumed as linear if
we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma 
\ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need 
to compute the optical force and torque applied to th dimer.

Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
or \textit{ott} \cite{Nieminen2007} we can compute the beam 
shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
type. Using the results from \cite{Farsund1996} we can then 
compute both the optical force and torque using the beam 
coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
which are found by calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
total force in along the z-direction and the total torque about 
the z-axis are provided:
\begin{equation}
	\label{eq:optical_force}
\begin{split}
  \bold{F_z}
  &=
    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
  \\ 
  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}
\begin{equation}
\begin{split}
  \bold{T_z}
  &=
    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
  \\ 
  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}

where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients
of the incident and scattered fields respectively. We can get the transverse 
force and torque components in a similar form by applying a simple rotation 
transformation. With the optical forces and torques computed all that remains
is to compute the Brownian forces and torques which are constrained by the relation.
\begin{align}
	\langle q_iq_j\rangle =2D_{ij}\Delta t
\end{align}

Vigilante et al compiled together a python package that combines both 
\textit{ott} and \textit{mstm} to simulate the behaviour of spherical 
aggregates within a predefined optical trap \cite{Vigilante2020}. Now 
\textit{ott} does have inbuilt $T$-matrix modules that can be used to 
model the scattering from different shaped particles; however, for 
aggregates of spheres the resultant force is nonsensical (see Chapter 
4 for further details). As such we rely on the \textit{mstm} for 
simulating dimers.

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

Secondly there is the particle frame which describes the orientation state of 
the particle, orientations are reported as quaternions which can be readily 
transformed into rotation matrices. The matrix describes the direction of the 
particles 3 orientation axis $U_x$, $U_y$, and $U_z$. The latter is oriented 
so that it is collinear with the centres of the two spheres; the direction of
vector $U_z$ is such that it goes from the centre of $a_{II}$ to the centre of
$a_{I}$. The origin of the 3 axis' is set on the particle's centre of mass, 
this is also used to define it's displacement from the origin of the laboratory 
frame. The diffusion tensor of the dimer is more complex than a sphere's as we 
must consider the rotational effects experienced by the dimer. We use the 
analytical solution of Nir \& Acrivos \cite{Nir1973}, where they provided the 
solution coefficients depending on the dimer's size ratio ($a_I/a_{II}$). Rather 
than recalculate the coefficients for every possible configuration of dimer we 
use a spline fitting function to approximate the diffusion tensor. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulated Quadrant Photodiode}
\label{sec:simulated_QPD}

Beyond merely simulating the forces experienced by a spherical aggregate we
also developed a simulative quadrant photo diode to replicate the results
from a typical calibration test. This builds upon the work from \cite{Rohrbach2002} which applied Lorenz-Mie theory to replicate the response signal of a QPD 
being used in back focal-plane interferometry. However, rather than be 
constrained to individual spheres we can now consider the expected response 
from any type of spherical aggregate.

In order to simulate a typical experimental set up with a QPD installed 
as a position detection system we need to evaluate the total magnitude of 
the electric field incident on the photo-diode surface. While trapping a 
micro-particle, the scattered and incident fields combine together and 
interfere with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD surface, the 
total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}

The last term is simply a representative step term that defines the outer 
limit by which we evaluate the electric field, this is analogous to our 
condenser lens removing noise from other light sources by only accepting 
light at a specific acceptance angle defined by its numerical aperture 
$NA_c$. Depending on the relative size of our particle we can adjust 
the acceptance angle, this has very little effect on the transverse 
signals, but for axial evaluations of a trapped particle the numerical 
aperture should be tuned so that the resultant response curve has negative 
slope in order to allow for axial position detection, the method for 
finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.

The incident beam is simple enough to define given our set up parameters, 
for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate. 
From the QPD's perspective it is receiving light both from the incident 
and scattered beam simultaneously, as such both fields must be expressed
using outgoing vector spherical harmonics.
\begin{align}
	E_{\rm inc}(r)&=
	E_0\sum^\infty_{nm}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
	\\
	E_{\rm scat}(r)&=E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
	-i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
	\right] 
\end{align}

We can, if we want, further restrict the incident beam by setting 
the truncation angle to match our microscope object, this applies a 
hard cut off to the incident beam to better represent experimental 
conditions. In order to compute the scattering from 
the target particle \textit{ott} uses the $T$-matrix method, this is 
not essential for a simple sphere but is essential for complex shaped 
particles such as dimers. The matrix from mstm needs to be repackaged 
to work with the \textit{ott} software, converting it from a column 
vector into a system of sub matrices.

The scattered and incident fields are then combined together 
in the far field to get $I(x,y)$, the quadrant and overall signals are 
calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}

Where the denominator is the total intensity on the QPD while there is 
no particle within the trap. The QPD sensitivity is dependent on both 
the polarisation of the incident beam and the displacement of the target 
sphere, this is shown in fig.~\ref{fig:totalfield}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fixed_polarisation.png}
	\captionsetup{margin=0.5cm}
	\caption{Total Field incident on the quadrant photo diode for a several
		different displacements. Top Left: $2\mu m $ diameter silica sphere
		placed at the focus of a  beam. Top Right: silica sphere is now
		displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere 
		is now displaced by $1 \mu m$ along the y-axis. Bottom Left: silica 
		sphere is now displaced by $0.707 \mu m$ along the x and y axis. 
		The trapping beam is a $TEM_{00}$ Gaussian, polarised along the 
		along the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of 
		each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
	\label{fig:totalfield}
\end{figure}

While translational motion have no knock on effects to the QPD signal, 
rotational changes are between the incident and total fields. This results
in rotational motion being biased in the QPD signal - even when collecting
signals from isotropic scatterers. To remedy this an inverse rotation is
applied to the total field.

To confirm that our method is producing accurate results, we ran a comparison
between our simulative QPD and the results from \cite{Rohrbach2002}. Where a 
$300\ nm$ diameter sphere is scanned across the path of a focused Gaussian beam
($\lambda=1064\ nm$, $NA=1.2$), the sphere has a refractive index of 1.57 and is
suspended in water ($n_{med}=1.33$) and the condenser lens has its numerical aperture
set to 0.5 ($\theta_{max} = 30^\circ$). Scanning across all three primary axis 
produced the following response curve:
\begin{figure}[h!]
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{QPD_axial_tests.png}
	\end{subfigure}
	\begin{subfigure}{0.475 \linewidth}
		\subcaption{}
		\includegraphics[width=\linewidth]{QPD_lat_tests.png}
	\end{subfigure}
	\caption{Comparison between QPD response signal versus work conducted by 
		Rohrbach, single sphere ($r = 150\ nm$, $n=1.57$) is scanned by a $1064\ nm$ laser and the QPD signal recorded. Solid lines represent the signal 
		produced by QPD using \textit{ott} and points represent the signal 
		response collected from \cite{Rohrbach2002}.}
	\label{fig:Rohrbach}
\end{figure}

The discrepancy between our simulated QPD and the results from 
Rohrbach can be attributed to the fact that the QPD signal  With this any trajectory can be collected from the QPD by displacing and 
rotating the beam accordingly, later in chapters \ref{chapter:langevin_dynamics} 
and \ref{chapter:simulated_detection} we discuss how to extract information 
from the QPD's signal.