### 2024-10-25 Latex_files/Thesis/main_win.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Strath PhD Thesis Template
%	by Jethro Browell [jethro.browell@strath.ac.uk]
%
%	Guidelines for thesis format, submission and content are found in
%	General and Course Regulations for Graduate and Postgraduate
%	Awards and Degrees, section 20.6.
%
%	Using .eps or .pdf is recomended to prduce high quality figures etc.
%
%	The Strathclyde logo can be found in other formats at www.strath.ac.uk.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,oneside,11pt]{book}
\setcounter{secnumdepth}{3}
\usepackage{amsbsy}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage[hidelinks]{hyperref}
%\usepackage{cite}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[strict]{changepage}
\usepackage{graphicx}
\graphicspath{{C:/Users/xbb16146/dmaciver97/Latex_files/Thesis/figs}}
\usepackage{sidecap}
% Page Margins - Strath Requirement
\usepackage[left=4cm,right=2.5cm,top=2cm,bottom=4cm,includehead,includefoot,headheight=15pt]{geometry}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage[
backend=biber,
bibstyle=ieee,
citestyle=ieee,
minbibnames=3, maxbibnames=3,
]{biblatex}
\addbibresource{thesis_bib.bib}

% Page Headers
\usepackage{fancyhdr}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % optional
%\fancyhead[L]{\nouppercase{\leftmark} \hfill Section \nouppercase{\rightmark}}
\fancyhead[L]{\nouppercase{\leftmark}}
\cfoot{\thepage}
\pagestyle{fancy}

% Draft Watermark
%\usepackage[draft=true,allpages=true,fontfamily=cmr,angle=90,scale=0.1,mark={\fboxsep=35pt\fboxrule=0pt\relax\fbox{-- DRAFT -- \today~--}},xcoord=-80,ycoord=-20]{draftmark}

\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand{\comment}[1]{{\bf\textcolor{red}{#1}}}
% \newcommand{\comment}[1]{{{#1}}}

% Line Spacing
%\def\baselinestretch{1.5} 
\usepackage{setspace}
\setstretch{1.5}

% Place UoS Logo on Title Page (this package modifies the "\maketitel" command.)
\usepackage{titling}
\pretitle{
  \begin{flushright}
  \vspace{-9.5cm}
%  \includegraphics[width=5cm,natwidth=472,natheight=531]{logo} \\[7cm]
  \includegraphics[width=5cm]{logo.pdf} \\[6cm]
  \end{flushright}
  \begin{center}
  \LARGE
  \posttitle{\end{center}}
  }

\title{Application of Micro-Rotors to Induce Crystal Nucleation: 
	   Challenges, Takeaways, and Future Prospects\\ PhD Thesis}
\author{D. Maciver\\[-0.8ex]
\small Chemical and Process Engineering Department\\[-0.8ex]
\small University of Strathclyde, Glasgow\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frontmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Declaration
\topskip0pt
\vspace*{\fill}
\noindent
\begin{quote}
	\centering
	This thesis is the result of the author's original research. It has been composed by the author and has not been previously submitted for examination which has led to the award of a degree. \\[5pt]
	%
	The copyright of this thesis belongs to the author under the terms of the United Kingdom Copyright Acts as qualified by University of Strathclyde Regulation 3.50. Due acknowledgement must always be made of the use of any material contained in, or derived from, this thesis. \\[5pt]
	%
	%Signed: \\
	%Date:
\end{quote}
\vspace*{\fill}

\chapter{Abstract}

\tableofcontents

\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{Preface/Acknowledgements}
I would like to acknowledge...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/chapter 1/main}

\input{chapters/chapter 2/main}

\input{chapters/chapter 3/main}

\input{chapters/chapter 4/main}

\input{chapters/chapter 5/main}

\chapter{Closing Remarks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\chapter{Appendix}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimation of fluid shear rates via galvano mirror}
An alternative approach to generating fluid shear is to use a 
galvano-mirror to rapidly move a trapped particle in a bulk 
fluid. While typically galvano and gimble mirrors are used to 
trap multiple particles simultaneously, a single micro sphere 
can be moved quickly through a fluid along a preset path. The 
only limitation on the particle's speed being the ratio of the 
trap stiffness to the drag force. For low Reynolds numbers we 
can assume that the fluid flow is inversely proportional to the
distance from the particle surface (see Eq.~\eqref{eq:birefringent_shear})

For a simple circular path one can estimate the sphere's speed 
by the radius of its path $R$ and the frequency of its orbit 
$\omega$ so that $U = R\omega$. For a more complex path, such 
as an elliptical orbit the curve needs to be parametrised. One 
can describe the position of a sphere on an elliptical path as 
such:
\begin{align}
	r(p) = \left[acos(2\pi p), bsin(2\pi p), 0 \right]
\end{align}

Where a and b are the characteristic radii of an ellipse. If we 
say that $p$ is some dimensionless parameter that describes the 
distance travelled along a trajectory (so $p=0$ is the start of 
the circular orbit and $p=1$ is just as we return to the start).
Then we can rewrite p as $p=t\omega$ where $\omega$ is the period
of orbit. Taking the partial derivative of the position gives 
us the sphere's velocity vector along that path. 
\begin{align}
	v(t) = \frac{\delta r(t)}{\delta t} = \left[-2\pi a\omega \ sin(2\pi t\omega),
	\ 2\pi b\omega \ cos(2\pi t\omega), 0 \right]
\end{align}

In order to compute the particle's speed $U$ we simply take the 
magnitude of our velocity. For low velocities the fluid flow at 
the sphere's surface can be computed based on its velocity.
\begin{align}
	u_r(r)=-|v(t)|\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $r$ is the radial distance to that point. Again taking the partial 
derivative we can get the shear rate for a particle moving through the fluid:
\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
	\label{eq:galvo_shear}
\end{align}

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable micro-rotors. Using 
\eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
equates to moving the silica bead around a circular path with a frequency of 
$100\ Hz$.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{galvano_shear_rate.png}
	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
		in bulk water moving at different speeds. The shear rate is calculated 
		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
		in a circular path and so the speed is constant through out its path.}
	\label{fig:galvano_shear}
\end{figure}

From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
mirror a better option for generating high shear rates but also over a larger 
volume. The only limitation to this method is the effective drag experienced 
by the trapped particle. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Bibliography}
\printbibliography

\end{document}