### 2024-11-18 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to chapter 4, among others

\chapter{Complex Langevin dynamics of spherical dimers}
\label{chapter:langevin_dynamics}
Much of the calibration theory discussed in Chapter 2 
assumes that the target particle in question is a single 
sphere, one whose scattering and motion is easily computed. 
However, while working with dense colloidal suspensions, 
one often ends up trapping more than one sphere. Li and 
Arlt \cite{Li2008} studied the case of two microspheres 
trapped in a single OT and found that multiple trapped 
beads could be mistaken for a single trapped bead with 
altered trap stiffness (see \ref{sec:harmonic_traps}). 
Theoretical studies on the case of two trapped microspheres 
by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
based model to show that the two trapped beads are brought 
into physical contact with each other by optical forces 
and they also calculated the axial equilibrium positions 
of the two trapped beads as a function of their size. 
Experiments in \cite{Praveen2016} confirmed that the two 
trapped beads indeed experience different trap stiffnesses 
in the vicinity of the same potential well.

There are further discussions looking into the dynamics of 
a whole host of asymmetrically shaped particles \cite{Loudet2014, 
ShengHua2005, Chetana2022}, their results all showing that 
predicting the behaviour an arbitrary shaped particle comes 
with great difficulty due to the fact that the optical force 
is dependent on a greater number of variables such as 
orientation and size factors.

With the initial goal of the project being to induce 
nucleation events via a rotating sphere, the aim of this 
chapter is to - in a limited capacity simulate and investigate 
the influence of a second particle being bound to our target 
sphere. The choice of a dimer, instead of an amorphous solid 
that might better represent a growing crystalline solid, 
allows us to consider how the dynamics of the aggregate change 
by varying a single parameter, namely the size ratio of the 
two spheres. Additionally, attempting to simulate an amorphous 
aggregate is rather difficult as calculating the optical force 
and torque is computationally slow. We build upon the works of 
Vigilante \textit{et al} \cite{Vigilante2020} to consider 
asymmetric spherical dimers and how varying size parameters 
alters the dynamics and additionally makes characterising their 
interactions within an optical trap more cumbersome.

In this chapter we will consider how the addition of a 
second sphere changes the trapping dynamics by introducing 
multiple equilibrium positions. These positions are 
dependent on both the size ratio of the dimer and its 
orientation relative to the trap. This is important to 
address, for while the the dimers' diffusive behaviour 
is the same, regardless of position and orientation, 
how it interacts with the trap may produce the wildly 
different results. Furthermore, we look at how our dimers 
interact with circularly polarised light, this is especially 
pertinent as the original plan was to utilise circularly 
polarised light to generate fluid flow in a supersaturated 
solution. The rotational motion of dimers in circularly 
polarised light goes against much of the established 
literature and suggests that the transfer of spin angular 
momentum is enabled by combining particles together. And 
lastly, we demonstrate how a quadrant photo diode, performs 
in characterising the actual interactions between a dimer 
and the optical trap. We do so to demonstrate that much 
of the dimer's trajectory information is either lost or 
poorly described by simple calibration techniques, making 
their motion difficult to characterise.

\subsection{Simulation Parameters}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
is focused by a objective with numerical aperture of 1.2 
and is x polarised. The size ratio of the two spheres is 
given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
\mu m$ unless specified otherwise; the dimer's orientation 
is given by a unit vector connecting the centres of both 
spheres (see \ref{sec:sim_parameters}). We define the 
'standard' orientation as being aligned with the direction 
of beam propagation direction - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation (see left hand side of 
figure~\ref{fig:paradigmatic}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
If we wanted to start from first principles and determine 
the trap strength on our target particle the first step 
would be to locate the equilibrium position relative to 
the trap focus. For a single sphere it is easy to enough 
to understand that its centre of mass will be drawn to 
focal point of the laser due to gradient forces, once there 
the force is analogous to a harmonic spring with a fixed 
trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
we consider instead a dimer, we now have two spheres both 
being drawn to the focus along by the same gradient force; 
in addition the scattering force is significantly more complex 
due to both spheres scattering the electromagnetic fields. 
This mutual scattering between individual spheres is what 
makes simulating spherical aggregates far more difficult 
compared to a single sphere, and even harder still to 
predict the position where the dimer's centre of mass is 
in equilibrium.

Because the scattering force is only significant in the 
direction of beam propagation \cite{Capitanio2002} we 
can assume there will only be a single potential well 
lying at the centre of the beam. Along the beam axis, 
such an assumption is no longer valid due to fact that 
spherical aggregates experience inter-particle scattering. 
This is a key reason for using \textit{mstm} as it 
accounts for that behaviour. The methodology for computing 
optical forces has been covered extensively for a number 
of different trapping conditions \cite{RanhaNeves2019}. 
So it is relatively easy to compute the trapping force 
and determine the equilibrium position by finding the 
position that minimises the net optical force and the 
local gradient is negative ($\delta F/\delta x < 0$) - 
we can assume that for a dielectric sphere the optical 
torque is negligible. For a dimer (or any arbitrary 
spherical aggregate), we now must consider both its 
position and orientation and find where the net optical 
force and torque are minimised.

After computing its $T$-matrix via \textit{mstm} and 
supplying that to \textit{ott} we compute the optical 
force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
in the axial direction while the dimer is in its 'standard' 
orientation. As expected we see a single point where 
the dimer will be in equilibrium, the linear fit in 
fig.~\ref{fig:paradigmatic}(a) indicates a that the 
force can be modelled as a harmonic potential close to 
the equilibrium position ($F_z\approx-\kappa_z z$). 
The second point where the axial force goes to 0 cannot 
be considered as equilibrium position as the positive 
gradient indicates that the trap is unstable unless 
Brownian motion is ignored.

We repeated the same calculation but now while the 
dimer is in its 'inverted' orientation, instead of 
a single point where the optical force is minimised 
we see that there are instead two separate equilibrium 
positions, one above the focus and one below the 
focus. In this particular example the two positions 
are far enough apart that both can be considered as 
separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the 
		centre of mass of the dimer (µm) for the 
		case of a dimer of size ratio 2. (a) is the 
		case where the dimer is in its' 'standard' 
		orientation, where the dimer is trapped at 
		$z = 1.71\ \mu m$. (b) is the case where the 
		dimer is in its' 'inverted' orientation, the 
		dimer is trapped at two positions: $z = 1.20\ 
		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
		renders to visualise the dimer orientation 
		are shown below each plot. The black lines 
		on each force-curve is a linear fit with the 
		slope being reported as the trap stiffness in 
		the legend.}
	\label{fig:paradigmatic}
\end{figure}

We can see that both equilibrium positions have comparable 
axial trap stiffness ($\kappa_z$), however the difference 
in the transverse trap stiffness ($\kappa_x$) is far more 
noticeable. The same dimer was trapped at each of the axial 
equilibrium positions and the transverse force was evaluated. 
While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable 
trap is far narrower in the 'standard' orientation compared 
to the 'inverted' cases. This highlights one of the challenges 
involved with studying asymmetric particles, even though its 
a simple enough process to trap them they maybe characterised 
very differently depending on their relative position and 
orientation towards the focus. This can have a significant 
impact on rheological studies - or attempting to probe any 
local property - as the variance in trap strength can result 
in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers
across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
for each trap. The same simulation parameters are used here 
as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
the dimer begins to approximate a single homogenous sphere 
- at least in terms of location and trap strength. However, 
for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found 
below the trapping focus. Previous work using the ray-optics 
model have confirmed even in the case that two spheres begin 
separated the electric field will align the particles as such 
that they make contact and are trapped together about a single 
trapping position \cite{Xu2005}. Furthermore it has been shown 
through proper manipulation of the Gaussian or Bessel beam 
modes that any number of trapping potentials can be developed 
\cite{Shahabadi2020} for nanoparticles. This result however, 
is the first example of an orientation dependent trapping 
situation using only a $TEM_00$ beam. Typical experimental 
arrangements cannot determine much information on the axial 
position of a trapped particle relative to the trap focus; 
this result indicates not only that dimers can be trapped 
in multiple axial positions but also their trapping behaviour 
is heavily dependent on said axial position. As such it is 
necessary that positional information in the z-axis can be 
elucidated if multiple spheres are trapped simultaneously. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned 
with the electric field is relatively simple as the orientational 
torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
trapped the dimer is unlikely to change orientation enough to 
escape the trap. Regardless, that does not rule out the 
possibility that there is a stable configuration where the 
orientation not strictly vertical, in fact most experimental 
work with symmetric nano-dimers will trap them lying perpendicular 
to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
\ref{sec:eq_positions} we cannot simply measure the optical force 
and torque as the parameter space is too large and determining if 
a particular position and orientation is stable is not clear based 
solely on force and torque measurements \cite{Bui2017}. Using the 
same simulation parameters as before we ran a number short 
simulations (total simulation time was $0.005$ s) with the laser 
power increased to $500\ mW$. Each simulation started with the 
dimer in a different starting position and orientation, due to the 
high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
between the direction of beam propagation and the dimer's orientation 
vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
divided into different regions depending on which equilibrium 
configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of
		size ratio 2 with a laser power of 500 mW ($\theta=0
		\circ$ is the 'standard' orientation and $\theta=180
		^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles 
		and the starting conditions are colour coded to match 
		the stable point they end up in. Right hand render 
		shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps 
are similar in magnitude to the vertically aligned dimers, 
but when the laser power is lowered (around $5\ mW$) the traps 
become metastable; after reaching its equilibrium configuration 
the particle behaves similarly to a typically trapped dimer but 
the trapping potential is small enough that the dimer escapes 
in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical 
configurations sees the dimer remaining trapped, even after 30 
seconds of run time, indicating that the trapping potential is 
far greater than the thermal energy. This suggests that the 
reason this off-axis configuration is metastable is due to the 
increased rotational freedom. Similar configurations have been 
explored with ellipsoids; Zhu \textit{et al.} looked at the 
dynamics of various elliptical particles and found that regardless 
of shape or initial orientation the particle would tend towards
either a purely vertical or purely horizontal orientation 
\cite{Zhu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion in circularly polarised light}
One aspect that has yet to be covered in depth with regards 
to spherical aggregates of any construction is their interaction 
with circularly polarised light. For homogenous spheres the 
optical torque is regarded as being negligible as the spin 
density cannot impart angular momentum while propagating in a 
homogenous medium. Dimers however, have been shown to experience 
an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations 
we found that dimers would rotate about their long axis when 
trapped in circularly polarised light. In this section we want 
to discuss how this behaviour is influenced by size, position, 
and orientation; and furthermore, we wish to address possible 
explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

\subsection{Polarisation Dependency on Dimer trajectory}
\label{sec:rot_pol}
In their paper Vigilante \textit{et al} attribute the rotational motion 
to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
scattering is a viable theory for describing optical rotation, however
there is usually an optical axis about which rotation occurs. For a dimer
this is typically the orientation vector that passes through the centres 
of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
cases where the particle's cross sectional shape is engineered to scatter
light in one particular direction \cite{Higurashi1994}, but if this was
the case then the dimer should rotate when illuminated by any polarisation
of light.

To that end, we simulated the motion of an optically trapped dimer in 
beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
orientational time series; the dimer's orientation is recorded as a 
quaternion which can be easily converted to a 3-dimensional rotation 
matrix. By considering only the transverse components ($U_{1,x}$, 
$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
the Fourier transformation of their time series reveals the rotational 
frequency. The laser power is set to $100\ mW$ to avoid large thermal 
fluctuations and so that the Fourier series of the transverse components 
approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
at the rotational frequency $\omega_{rot}$.

%% Example plot of rotating beam trajectory and then the same thing but 
%% Fourier transformed
\begin{equation}
	\begin{split}
		q(t) \rightarrow R(t) =& 
		\begin{pmatrix}
			U_{1,x}(t) && U_{2,x}(t) && U_{3,x}(t) \\
			U_{1,y}(t) && U_{2,y}(t) && U_{3,y}(t) \\
			U_{1,z}(t) && U_{2,z}(t) && U_{3,z}(t) 
		\end{pmatrix} \\
		\rightarrow
		&\int^\infty_{-\infty}R(t)e^{-i2\pi ft} dt = 
		\begin{pmatrix}
			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f) \\
			\delta(\omega_{rot}-f) && \delta(\omega_{rot}-f) && \delta(f)\\
			\delta(f) && \delta(f) && \delta(f)
		\end{pmatrix}
	\end{split}
\end{equation}

If the rotational frequency was not immediately obvious the simulation was
repeated but over a longer simulation time. Four different size ratio of 
dimers were studied, both in their 'standard' and 'inverted' orientations. 
The results of this are displayed in Fig.~\ref{fig:rotation_vs_pol}:
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{rotation_vs_pol.png}
	\caption{Rotation frequency vs component phase difference for 
		differently sized dimers. The solid lines represent the 
		rotation rate experienced while the dimer is in its standard 
		orientation, whereas the solid points are for the case 
		where the orientation is inverted. Laser power = $100\ mW$.}
	\label{fig:rotation_vs_pol}
\end{figure}

This shows us that these optical rotations are polarisation 
dependent and not merely an example of the dimer scattering 
light asymmetrically. We also know that this cannot be a 
similar mechanism to that of experimental works \cite{Ahn2018, 
Reimann2018} as in our simulations the dimer's long axis
is parallel to the electric field. In this orientation 
the dimer's optical axis is aligned perpendicular to the
polarisation vector meaning that there should be no angular
momentum transferred along the optical axis. This is consistent 
with other experiments involving elongated particles. A dual
beam trap was used to study the dynamics of "disk like" particles. 
They found that these particles had one of two stable
orientations: Either orienting with the 'flat' side perpendicular
to the beam trap, with no rotation being observed \cite{Brzobohaty2015}. 
Or orientating with the 'flat' side parallel to the trap, in which 
case the long axis was aligned with the polarisation vector and 
thus rotational motion was detected \cite{Brzobohaty2015}.

\subsection{Brownian Vortex via Curl of Spin momentum}
In their work Vigilante refers to the spin-curl effects 
demonstrated by Grier \textit{et al} \cite{Ruffner2012}, 
in which the curl of the spin density leads to a second 
order optical force that orbits around the beams central 
axis \cite{Yevick2017}. While several papers have 
demonstrated this phenomena \cite{Zhao2007, Zhao2009, 
Wang2010} it was only properly formalised by \cite{Ruffner2012}. 
In which they showed that the seemingly random trajectory 
of a trapped sphere was biased by the polarisation state 
of the trapping beam. While not immediately evident from 
the trajectory the helicity of the trapping beam was revealed 
by computing the particle's probability flux using.
\begin{align}
	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
	\label{eq:prob_flux}
\end{align}

where $\delta_{\sigma_j}$ is the kernel of an adaptive 
density estimator \cite{Silverman1986}. \eqref{eq:prob_flux} 
describes the direction a trapped sphere is most likely 
move in given our statistical knowledge of the 
trajectories probability density function. A finite 
estimation of the density function $p(r)$ is used in 
\cite{Ruffner2012}.
\begin{align}
	p(r) = \frac{1}{N}\sum_{j=1}^N \delta_{\sigma_j}(r-r_j)
	\label{eq:prob_density}
\end{align}

The probability flux reveals a biased motion in the trajectory 
of a single sphere (see Fig.~\ref{fig:Ruffner_Grier}). This 
biased motion results in a slight orbital motion about the 
central axis of the trapping beam, the orbital frequency is 
shown to be proportional to the polarisation state of the 
trapping beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how 
		the momentum density of a Gaussian beam is twisted while 
		using circularly polarised light. The top row (figures 
		(b)-(d)) shows a 7 sphere cluster trapped in a circularly 
		polarised beam. Due to the clusters asymmetric susceptibility 
		to polarization the cluster rotates in the $x-y$ plane. 
		Whereas the bottom row (figures (e) - (g)) show the similar 
		results for a single sphere. In this instance the sphere 
		does not rotate but instead orbits the beam axis. In both 
		instances the motion is proportional to the degree of 
		polarisation (see figures (d) and (g)) but for the single 
		sphere this motion is only revealed when using 
		\eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused 
		with permission from author}
	\label{fig:Ruffner_Grier}
\end{figure}

While the results from \cite{Ruffner2012} suggest that the 
optical rotation seen in asymmetric dimers can be attributed to 
the same spin-curl forces there are several questions that 
cannot be explained purely by the spin-curl force.

\subsection{Optical torque differences}
In their paper Grier \textit{et al} show that a 7 sphere cluster
will rotate when trapped in circularly polarised light but a 
homogenous sphere will have a slight curl to its trajectory. 
They attribute this to the curl of the spin angular momentum
generating a "Brownian vortex". But what is more interesting, 
and something that appears unaddressed, is how the effects of 
this vortex change based on the overall shape of the target
particle.

You would not expect that the addition of a single additional 
sphere should drastically adjust the torque especially if said 
sphere is relatively small. However when we measured the optical 
torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we 
found the exact opposite. In both cases we used the same trapping 
beam as used for fig.~\ref{fig:rotation_vs_pol} but with a 
circularly polarised beam. Both the sphere and dimer were rotated 
in the $x-z$ plane and the all three components of the optical 
torque were recorded.
\begin{figure}[h!]
	\centering	
	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
	\caption{Optical toque experienced by a dimer ($a_{I}/a
	{II}=10$) and a single isotropic sphere. Both were rotated 
	in the $x-z$ plane and the angle between $U_z$ and the beam 
	axis gives the polar angle.The solid lines denote the torque 
	experienced by the dimer whereas the dashed lines represent 
	the torque experienced by the sphere.}
	\label{fig:sphere_dimer_torque}
\end{figure}

The torques about the $x$ and $y$ axis can be somewhat understood 
as the second sphere is being drawn back towards the centre of the 
trap by the gradient forces. The same cannot be said for the $z$-
component of the optical torque, which is non-zero even at a 
$80^\circ$ angle. This cannot be simply explained via spin-curl 
effects, as the 'Brownian vortex' should instead be driving the
dimer around the beam axis - similar to how the 7 sphere cluster 
behaved. What is clear that the mere combination of two spheres
results in a unique behaviour that has not been previously 
investigated.

\subsection{Rotational frequency as a function of size ratio and 
	orientation}
\label{sec:rot_size}	
One factor t Intuitively, you would expect that a 
larger particle would experience a greater torque and therefore 
rotate faster. By repeating the same kinds of simulation as used 
in \ref{fig:rotation_vs_pol} but for a circularly polarised beam 
$\phi=90\circ$ it was found that not only is the rotation rate 
dependent on the size of the dimer, but also on its orientation 
and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while 
  	trapped in a circularly polarised beam; a positive rotation 
  	rate indicates clockwise rotation, whereas a negative 
  	rotation rate indicates counter-clockwise rotation. The red 
  	line is for the case of a dimer in its 'standard' orientation. 
  	The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the 
  	beam. And lastly the green line is for the case when the 
  	dimer is in its 'inverted' orientation, but when it is trapped 
  	below the focus of the beam.}
\end{figure}

It is difficult to see from the graph, but the rotation rate never
truly goes down to zero, reaching a minimum of $2\,{\rm Hz}$, which 
would imply that a second sphere of radius $200\,{\rm nm}$ is enough 
to induce rotational motion. What is also interesting is that the 
rotation rate is not correlated directly with either the particle 
size or the equilibrium position. This is in stark contrast to previous
reports of dimer optical rotation; Ahn \textit{et al.} reported that
the rotational frequency is maximised when the dimer is symmetric 
\cite{Ahn2018}. Their work was conducted in a vacuum so the only 
limiting factor to the optical rotational motion is the structural 
stability of the particle itself.

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity sized 
dimers there is the potential for non-vertical trapping orientations 
in which the dimer is still stably trapped. When a circularly 
polarised beam is used the dimer exhibits gyroscopic precession. 
As shown in fig.~\ref{fig:gyro} the dimer's trajectory exhibits 
periodic rotation, both around its long axis and about the beam 
axis.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped 
		in an off axis orientation with a circularly polarised beam 
		($P= 100\ mW$). The far left column depicts the dimer's centre 
		of mass position with time; the remaining 3 columns show the 9
		components of the dimers' rotation matrix, with each column being
		associated with one of its three principal axis. }
	\label{fig:gyro}
\end{figure}

Applying a Fourier analysis to the above trajectory reveals 
the 3 fundamental frequencies typically associated with 
gyroscopic precession. Fig.~\ref{fig:gyro_diagram} gives 
a representative idea of the motion seen in fig.~\ref{fig:gyro}.
Firstly there is precession, is denoted by $\psi$, this
causes the dimer to rotate in the $XY$ plane while maintaining
a constant polar angle. The precession can be due to the dimer 
having inhomogeneous polarisation susceptibility. Since its 
long axis is more susceptible to being polarised the dimer will
try to align its long axis with the polarisation vector 
\cite{Bruce2020}. Second there is nutation, denoted by $\theta$, 
which is due to the dimer's Brownian motion being influenced by 
the 'Brownian vortex', as predicted by \cite{Ruffner2012}. And 
lastly there is rotation, denoted by $\omega$, where the dimer
spins around its long axis as it did in \ref{sec:rot_pol} \&
\ref{sec:rot_size}.

What is interesting is that while we see the precession and 
nutation also contribute to the rotation frequency. As can be 
seen in both fig.~\ref{fig:gyro} and fig.~\ref{fig:gyro_diagram}
the rotation matrix components $U_{x,x}$, $U_{x,y}$, $U_{y,x}$
and $U_{y,y}$

\begin{figure}
	\hspace{-2.3cm}
	\includegraphics[width=1.2\linewidth]{gyroscopic_diagram.pdf}
	\caption{Representative diagram of the gyroscopic precession 
	from fig.~\ref{fig:gyro}. The dimer has three principal angular
	frequencies: The rotation ($\omega$) occurs around the 
	the dimer's long axis. Precession ($\psi$) is seen where the dimer 
	rotates around the beam axis. Nutation ($\theta$) is due to the dimer's 
	centre of diffusion orbiting the beam axis. Shown on the right
	is the power spectra from fig.~\ref{fig:gyro} with the associated 
	frequencies labelled. The power spectrum have been zoomed in on
	the relevant frequencies to highlight the precise values.}
	\label{fig:gyro_diagram}
\end{figure}

This gyroscopic motion has been demonstrated previously in 
nanoparticles \cite{Zhu2021, Rashid2018, Hoang2016, Kuhn2016} 
but has not been observed for micron scale aggregates. What
is interesting that the precession is seen around the particle's
optical axis (in the case of a dimer this would be its long 
axis). Fig.~\ref{fig:gyro} instead shows the precession 
occurring perpendicular to the optical axis. In the former 
case the precession can be explained as simply occurring due 
to the reaction torque from the surrounding fluid \cite{Zhu2021}.
In their analysis of ellipsoidal particles, Zhu \textit{et al.}
found that when the particle rotated around their minimal inertial 
axis (similar to the observed behaviour in dimers), the particle 
would quickly 'stabilise' by aligning its' optical axis with the 
beam axis. This highlights that a spherical dimer is capable of 
rotating about its minimal inertial axis while remaining stably
trapped. This has potential implications beyond simply 
understanding particle dynamics.

Further analysis of the mechanism behind the precession of off-axis 
dimers may provide insights into controlling Brownian motion. An 
experimental work trying to 'cool' nano-dimers by controlling the 
motion in all 6 degrees of freedom found that even while the 
rotation about the short axis' could be controlled the free rotation 
around the dimers' long axis resulted in an unpredictable torsional 
vibration \cite{Bang2020}. Understanding how rotational motion 
arises in the Mie-regime could allow researchers to build a robust 
theoretical framework to construct beam structures that eliminate any 
unwanted rotational motion from a target particle. Conversely, the 
same framework could allow for precise measurements of the optical 
torque applied to a target particle, allowing for characterisation 
of complex shaped particles' interactions with an optical trap. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterisation of asymmetric dimers via PSD \\
	analysis}
As discussed in \ref{sec:simulated_QPD}, one of the methods 
developed to work in conjunction with \cite{Vigilante2020} 
is a simulated quadrant photo diode for as a position 
detection system. While it is possible to extract all of 
the relevant dynamical information from a simulation, 
confirming the same behaviour in an experimental setting 
can be challenging if dealing with a non-birefringent 
anisotropic scatterer. The QPD is composed of 4 photodiodes 
that measure the intensity of light incident on their 
surfaces. Using \textit{ott} we can define a region in the 
far-field that is analogues to the surface of the QPD surface. 
Evaluating the electric field across this surface gives us 
an approximation of the QPD signal outputted in an experimental 
situation. This does not provide a one-to-one result however 
as hardware errors (i.e. internal resistance, external light 
sources, and vibrations) would distort the signal somewhat. We 
can use this as a means to examine the limitations of using 
back-focal plane interferometry as a means for characterising 
anisotropic scatterers.

As a benchmark we start by considering a single sphere within 
an optical trap. A single polystyrene sphere suspended in water 
($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
Gaussian beam ($NA=1.25$) using circularly polarised light. For 
the sake of time efficiency the trajectory was sampled every 10
time steps, meaning the upper bound on the power spectra is
$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
we fitted the power spectra using the aliased Lorentzian 
(Eq.~\eqref{eq:aliased_lorentzian}).
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~
		\ref{eq:alaised_lorentzian}, scattered points represents 
		the blocked data ($n_b=100$). Corner frequency for the 
		Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
report different corner frequencies which would be indicate 
that the trap is not perfectly circular. We can use both 
\textit{ott} and the trajectory itself to derive an estimation 
of the trap geometry. The corner frequencies and corresponding 
trap stiffness are reported below:

\begin{center}
	\captionof{table}{QPD fitting for single sphere}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
		fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline

	\end{tabular}
\end{center}

Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
where the variance is based upon our choice of frequency 
boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
the ellipticity of the beam is given by $e = (1-\kappa_y/
\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
wavefront. Its clear from these initial results that the QPD 
is more sensitive to changes along the y-axis than the x-axis 
when compared to the direct \textit{ott} calculations. This is 
somewhat reflected in the trajectory results. Typically, even 
an industrial Gaussian beam will produce an elliptical diffraction 
limit spot when heavily focused; in their tutorial for optimizing 
the PSD analysis, Berg and Sorensen reported a ellipticity of 
around 15 \% after a total calibration time of 80 seconds 
\cite{BergSoerensen2004}.

The reason for the discrepancies between all 3 methods is due 
to what is actually being measured. The \textit{ott} estimates 
are simply looking at the differences in the trapping strength 
along the Cartesian axis'. If calibrated over an long enough 
time frame you would expect that the resulting power spectra 
would exactly mirror the \textit{ott} predictions. However over 
a short calibration time a QPD will only sample a small trajectory 
that does not fully explore the trapping plane. There is a clear 
trade off in terms of accuracy and computation time as shorter 
calibration runs are computationally more efficient but prone to 
errors. This is the case even if we know the exact positional data.

With this in mind, let us consider a symmetric dimer that is 
optically trapped by the same Gaussian beam. Not only does 
the dimer's equilibrium position change but it is subjected 
to rotational motion due to its unequal moments of inertia. 
This is reflected in the calibration results using the simulated 
QPD, where we see a drastically different estimation between the 
\textit{ott} estimate and the QPD estimate.

\begin{center}
	\captionof{table}{QPD fitting for symmetric dimer}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Simulation fitting} \\
		\hline
		$f_c\ [Hz]$ & 409 & 334 & \parbox{1cm}{\centering 431} & 424 
		& \parbox{1.25cm}{\centering 274} & 285 \\
		$\kappa\ [pN/\mu m]$ & 48.51 & 39.58 & 51.13 & 50.26 & 32.45 & 33.75 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{42.8 \%} &
		\multicolumn{2}{|c|}{12.7 \%} & 
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

Now we see that the \textit{ott} predicts a more elliptical 
trap compared to the QPD model which says the trap is far
more symmetrical while trapping a symmetric dimer. A potential 
reason that \textit{ott} no longer expects a circular trap 
could be due to how it computes the beam shape coefficients; 
by point matching in the far field before the focus means a 
loss in accuracy for objects that trap above the focus. 
The change in the QPD estimation can be partially explained 
by the fact that rotational effects are not accounted for in 
the Lorentzian power spectra, only translational motion. 
Typically, rotational motion is only ever detected when it 
is periodic (take for example Fig.~\ref{fig:vaterite}), when 
the motion is stochastic the entire power spectra is effected 
making it near impossible to separate the translational and 
rotational contributions from a single power spectra.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
Considering the simplicity of a scatterer such as a dimer, one would 
assume that the dynamics of such an object would be relatively easy 
to predict. Simulations of dimers in the Mie regime show that not 
only do they have multiple positions and orientations in which they 
can be trapped but also that their interaction with circularly 
polarised light is heavily dependent on the axial position and 
trapping orientation.

Dimer's have the potential to be used as tunable micro-rotors, being 
simple to synthesise and can be made out of any material of choice.
The rotation demonstrated by dimer's is not accurately described in
previous literature which raises questions on the interactions between
circularly polarised light and spherical aggregates. Until now, our 
understanding of angular momentum transfer has been either focused on 
single particles (where momentum transfer is easily described by 
Lorenz-Mie theory) or for large aggregates of particles, where the 
angular momentum transfer between individual particles is not considered. 
This could provide a greater understanding of angular momentum and help in 
the development of better torque sensing or torque preventing methods.

Conventional calibration methods will not be adequate however, while
angular motion can be detected using complicated set ups \cite{Bang2020},
this can only provide an estimate of the magnitude of a particle's 
angular motion. This is perfectly fine in cases where the rotational
motion is stochastic and intuitively predictable (i.e. a dimer in a
vertical orientation undergoing stochastic Brownian motion), but in 
the case the the motion is instead periodic characterising it requires
instantaneous measurements of the particle's motion.